---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.path = "man/figures/README-",
  out.width = "80%"
)
```

# optedr

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/optedr)](https://CRAN.R-project.org/package=optedr)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

## Overview

**optedr** is an optimal experimental design suite for non-linear and multi-factor models. Its main capabilities are:

- Calculating **D-, Ds-, A-, I- and L-optimal** designs for non-linear models with any number of factors.
- Calculating **Compound** designs combining multiple criteria with user-specified weights.
- Calculating **KL-Optimality** designs for model discrimination, supporting Normal, Poisson, Binomial and Gamma families.
- **Augmenting** existing designs with controlled D-efficiency loss.
- Evaluating the **efficiency** of a design against the optimum.
- **Rounding** approximate designs to exact designs.

## Installation

You can install the released version of optedr from [CRAN](https://CRAN.R-project.org) with:

```{r, eval=FALSE}
install.packages("optedr")
```

You can install the latest version from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
devtools::install_github("kezrael/optedr")
```

## Functions

- `opt_des()` — calculates optimal designs.
- `design_efficiency()` — evaluates the efficiency of a design against the optimum.
- `get_augment_region()` — computes the candidate region for augmentation with controlled efficiency loss.
- `augment_design()` — adds points to a design and rescales weights.
- `efficient_round()` — rounds approximate designs to exact designs (multiplier method).
- `combinatorial_round()` — rounds approximate designs to exact designs (exhaustive floor/ceiling search).
- `make_kl_fun()` — constructs a KL divergence function from two model-family pairs.
- `shiny_optimal()` / `shiny_augment()` — interactive Shiny demos.

The `optdes` object returned by `opt_des()` has its own `print()`, `summary()` and `plot()` methods.

## Usage

```{r, eval=FALSE}
library(optedr)
```

```{r, echo=FALSE}
devtools::load_all()
```

### Optimal design

`opt_des()` requires a `criterion`, a `model`, `parameters` and their nominal values, and a `design_space`. For multi-factor models pass `design_space` as a named list — variable names must match those in `model`.

```{r opt-des}
library(optedr)

result_D <- opt_des(
  criterion    = "D-Optimality",
  model        = y ~ Vmax * x1 * x2 / ((K1 + x1) * (K2 + x2)),
  parameters   = c("Vmax", "K1", "K2"),
  par_values   = c(1, 1, 1),
  design_space = list(x1 = c(0.1, 10), x2 = c(0.1, 10))
)
result_D
```

`plot()` returns a heatmap of the sensitivity function, with support points overlaid in red. By the Equivalence Theorem the design is optimal if and only if the sensitivity equals the number of parameters at every support point and is below it everywhere else.

```{r opt-plot}
plot(result_D)
```

`summary()` gives additional detail: criterion value, Atwood convergence criterion, and the design space.

### Other criteria

The same interface supports five optimality criteria:

| Criterion | Key argument | Objective |
|-----------|-------------|-----------|
| `"D-Optimality"` | — | Minimise $-\log\det M$ |
| `"Ds-Optimality"` | `par_int` | Focus on a subset of parameters |
| `"A-Optimality"` | — | Minimise $\operatorname{tr} M^{-1}$ |
| `"I-Optimality"` | `reg_int` | Minimise integrated prediction variance |
| `"L-Optimality"` | `matB` | Minimise $\operatorname{tr}(B\,M^{-1})$ |

For example, A-Optimality minimises the average variance of all estimators:

```{r a-opt}
result_A <- opt_des(
  criterion    = "A-Optimality",
  model        = y ~ Vmax * x1 * x2 / ((K1 + x1) * (K2 + x2)),
  parameters   = c("Vmax", "K1", "K2"),
  par_values   = c(1, 1, 1),
  design_space = list(x1 = c(0.1, 10), x2 = c(0.1, 10))
)
result_A
```

### Compound criterion

Combine any set of criteria as a weighted sum. Weights are normalised automatically.

```{r compound}
result_DI <- opt_des(
  criterion    = "Compound",
  model        = y ~ Vmax * x1 * x2 / ((K1 + x1) * (K2 + x2)),
  parameters   = c("Vmax", "K1", "K2"),
  par_values   = c(1, 1, 1),
  design_space = list(x1 = c(0.1, 10), x2 = c(0.1, 10)),
  compound     = list(
    list(criterion = "D-Optimality", weight = 0.6),
    list(criterion = "I-Optimality", weight = 0.4,
         reg_int = list(x1 = c(1, 5), x2 = c(1, 5)))
  )
)
result_DI
```

### Design efficiency

`design_efficiency()` measures the fraction of the optimal information that a given design achieves. An efficiency of 0.80 means the design needs $1/0.80 = 1.25\times$ more observations to match the optimal.

```{r efficiency}
corners <- data.frame(
  x1     = c(0.1, 10,  0.1, 10),
  x2     = c(0.1, 0.1, 10,  10),
  Weight = rep(0.25, 4)
)
eff <- design_efficiency(corners, result_D)
```

### Augmenting designs

`get_augment_region()` finds which points can be added to an existing design while keeping its D-efficiency above a threshold `delta_val`. `augment_design()` then adds the chosen point and rescales weights. Both functions work in 1D and multi-dimensional spaces.

```{r augment, message=FALSE, results='hide'}
init_des <- data.frame(
  x1     = c(0.8, 10,  5),
  x2     = c(10,  0.8, 5),
  Weight = c(1/3, 1/3, 1/3)
)

region <- get_augment_region(
  criterion           = "D-Optimality",
  init_design         = init_des,
  alpha               = 0.25,
  model               = y ~ Vmax * x1 * x2 / ((K1 + x1) * (K2 + x2)),
  parameters          = c("Vmax", "K1", "K2"),
  par_values          = c(1, 1, 1),
  design_space        = list(x1 = c(0.1, 10), x2 = c(0.1, 10)),
  calc_optimal_design = FALSE,
  delta_val           = 0.85
)
```

`region$region` contains the candidate points ranked by efficiency gain. Pick the best and augment:

```{r augment-add, fig.show='hide'}
best <- region$region[which.max(region$region$efficiency), ]

augmented <- augment_design(
  criterion           = "D-Optimality",
  init_design         = init_des,
  alpha               = 0.25,
  model               = y ~ Vmax * x1 * x2 / ((K1 + x1) * (K2 + x2)),
  parameters          = c("Vmax", "K1", "K2"),
  par_values          = c(1, 1, 1),
  design_space        = list(x1 = c(0.1, 10), x2 = c(0.1, 10)),
  calc_optimal_design = FALSE,
  delta_val           = 0.85,
  new_points          = data.frame(x1 = best$x1, x2 = best$x2, Weight = 1)
)
augmented
```

### Rounding to exact designs

Approximate designs assign continuous weights. Two functions convert them to integer replication counts for a given total $n$:

```{r round}
# Multiplier method
efficient_round(result_D$optdes, n = 12)

# Exhaustive floor/ceiling search (optimal for small number of support points)
combinatorial_round(result_D, n = 12)
```

### KL-Optimality

KL-Optimality finds the design that maximises the Kullback-Leibler divergence between a reference model and the most adversarial rival, for Normal, Poisson, Binomial and Gamma families.

```{r kl}
result_kl <- opt_des(
  "KL-Optimality",
  model        = y ~ Vmax * x / (Km + x),
  parameters   = c("Vmax", "Km"),
  par_values   = c(2, 1),
  design_space = c(0.1, 5),
  rival_model  = y ~ a * x,
  rival_params = c("a"),
  rival_pars   = c(1),
  rival_lower  = c(0.2),
  rival_upper  = c(2.5),
  family       = "Normal",
  phi          = 1
)
result_kl
plot(result_kl)
```

`make_kl_fun()` automates the KL function construction for custom model-family pairs, including different error structures or families.

## Learn more

- `vignette("optedr-intro")` — optimality criteria, multi-dimensional design spaces, compound designs, efficiency and rounding.
- `vignette("optedr-augment")` — augmenting designs in 1D, 2D and 3D with controlled efficiency loss.
- `vignette("optedr-kl")` — KL-Optimality for model discrimination, `make_kl_fun()`, and the Hill model case study.
